A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution

### 2018-06-13 paper/paper.tex
Unifying a bit the presentation between the main part and the appendix

In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to be placed on grid coordinates while avoiding vertices to overlap with other vertices or edges. Further, we restricted the grid size to \begin{inparaenum}[(i)]
\item $10^6 \times 10^6$
\item $10^4 \times 10^4$
\item $10^3 \times 10^3$, and
\item $10^2 \times 10^2$.
\end{inparaenum} The test set for this experiment was again the set of non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, but we computed a different initial since our algorithm requires its input drawing to maintain its invariants (that is, vertices must be on the grid). More precisely, we computed a random grid layout of each graph where vertices to overlap with other vertices or edges. For each grid size, we computed a layout with the variant of our algorithm focusing on crossing resolution. Fig.~\ref{fig:experimentsGrid} summarizes the results.

@@

\subfloat[\label{fig:crossingResolutionCrossingNumberGrid}{No.~of crossings vs no.~of vertices}]{
\centering
\scalebox{0.95}{
\centering
%======================================
% Crossing Resolution Crossing Number
%======================================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
xmin=10,
xmax=100,
ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
\addplot table [x=n, y=Crossing number rm1M, col sep=semicolon]{gridExperiment.csv};
\addplot table [x=n, y=Crossing number rm10k, col sep=semicolon] {gridExperiment.csv};
\addplot table [x=n, y=Crossing number rm1000, col sep=semicolon]{gridExperiment.csv};
\addplot table [x=n, y=Crossing number rm100, col sep=semicolon] {gridExperiment.csv};
\addplot table [x=n, y=Crossing number rm, col sep=semicolon] {gridExperiment.csv};
\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
\end{axis}
\end{tikzpicture}}}
\subfloat[\label{fig:crossingResolutionTimeGrid}{Iterations vs no.~of vertices}]{
\centering
\scalebox{0.95}{
\centering
%==================================
% Crossing Resolution Iterations
%==================================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
xmin=10,
xmax=100,
ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left, 
ylabel={Iterations},
xlabel={Number of Vertices},ymajorgrids]
\addplot table [x=n, y=Iterations rm1M, col sep=semicolon]{gridExperiment.csv};
\addplot table [x=n, y=Iterations rm10k, col sep=semicolon] {gridExperiment.csv};
\addplot table [x=n, y=Iterations rm1000, col sep=semicolon]{gridExperiment.csv};
\addplot table [x=n, y=Iterations rm100, col sep=semicolon] {gridExperiment.csv};
\addplot table [x=n, y=Iterations rm, col sep=semicolon] {gridExperiment.csv};
\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
\end{axis}
\end{tikzpicture}}}
\caption{Illustration of our experimental results on the crossing resolution with grid restriction. The double line shows the results of our unrestricted algorithm from Section~\ref{sec:experiments}.}
\label{fig:experimentsGrid}
\end{figure}

Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}. More precisely, the grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of graphs which often resulted in the initial layout to be accepted as the best solution for larger graphs. Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}). For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout. For grid size $10^4 \times 10^4$, the algorithm performed only about $10^\circ$ worse than the unrestricted version of our algorithm which could only slightly be improved by increasing the grid size to $10^6 \times 10^6$.

The aspect ratio of computed drawings was more or less the same for all grid size with the exception of size $10^2 \times 10^2$ and always about twice the aspect ratio of the unrestricted variant of our algorithm; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}. For grid size $10^2 \times 10^2$, this may again be explained by the fact, that for the majority of graphs, the initial layout could not be improved.

The number of crossings increased with the restriction on the size of the grid; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}. As with the crossing resolution, there are clear differences between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$ whereas there is only a slight improvement from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$ which still uses twice as many crossings as our unrestricted variant. This can partially be explained with the different choice of the initial drawing.

Finally, we observe that the number of iterations needed for convergence for grid size $10^3 \times 10^3$ was already twice as much as for our unrestricted algorithm; see Fig.~\ref{fig:crossingResolutionTimeGrid}. From there on, the number of iterations for convergence increases with the grid size. Note that grid size $10^6 \times 10^6$ needed about 1000 to 2000 iterations more than grid size $10^4 \times 10^4$ even though the final layout was only marginally better as discussed prior. The curve for $10^2 \times 10^2$ highlights that this grid size was too restrictive for the vast majority of graphs which often resulted in the initial layout to be accepted as the best solution for larger graphs.

In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid as long as the grid is not too small. However, the computation of grid drawings takes much longer than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order to improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution. Further experiments in this direction are needed.

@@

We repeated our experimental evaluation on the crossing resolution, total resolution and angular resolution without grid constraint on a second test set of graphs, the set of non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}). The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}. In general we observed that the variance of results is much larger than in the experiments on the Rome graphs. This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of graphs in this second test set varies vastly between different graph sizes.

@@

For the crossing resolution we observed that both the unrestricted and the aspect ratio restricted variants of our algorithm again outperformed the two evaluated force directed approachs; see Fig.~\ref{fig:northCrossing-1}. Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution on different graph sizes as the curves are nearly parallel. By all these results, we can classify the graphs into ``hard'' or ``easy'' graphs regarding the crossing resolution maximization; in particular graphs between 50 and 70 vertices appear to be harder to improve than graphs between 70 and 80 vertices. Regarding the aspect ratio of the produced drawings we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force directed algorithms appears to be quite unstable resulting in a large variance. Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio which is much lower than the one of our unrestricted algorithm for larger graphs. All four algorithms behave nearly the same in terms of number of crossings; see Fig.~\ref{fig:northCrossing-3}. In terms of the number of iterations, we observe that somewhat surprisingly the algorithm of Argyriou et al. converges in the least amout of iterations throughout the test set whereas the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.

\begin{figure}[htb]
\centering
\subfloat[\label{fig:northTotal-1}{Total resolution vs no.~of vertices}]{
\centering
\scalebox{.95}{
\centering
%======================
% Total Resolution
%======================
\begin{tikzpicture}
\begin{axis}[
axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
width=0.475\textwidth,
cycle list name=mylist, 
mark repeat={5},xlabel={Number of Vertices},ylabel={Total Resolution},
xtick={10,20,30,40,50,60,70,80,90,100},
xmin=10,
xmax=100,
ytick={10,20,30,40,50,60,70,80,90},
tick pos=left,ymajorgrids]
\addplot table [x=avg, y=tr rm, col sep=semicolon]{totalResolutionNorth.csv};
\addplot table [x=avg, y=tr rm-1, col sep=semicolon] {totalResolutionNorth.csv};
\addplot table [x=avg, y=tr fa, col sep=semicolon]{totalResolutionNorth.csv};
\addplot table [x=avg, y=tr tr, col sep=semicolon] {totalResolutionNorth.csv};
\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
\end{axis}
\end{tikzpicture}}}
\subfloat[\label{fig:northTotal-2}{Aspect ratio vs no.~of vertices}]{
\centering
\scalebox{.95}{
\centering
%===================================
% Total Resolution Aspect Ratio
%===================================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
xlabel={Number of Vertices},
xmin=10,
xmax=100,
ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
\addplot table [x=avg, y=ar rm, col sep=semicolon]{totalResolutionNorth.csv};
\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {totalResolutionNorth.csv};
\addplot table [x=avg, y=ar fa, col sep=semicolon]{totalResolutionNorth.csv};
\addplot table [x=avg, y=ar tr, col sep=semicolon] {totalResolutionNorth.csv};
\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
\end{axis}
\end{tikzpicture}}}

@@

In the total resolution experiment, for smaller graphs,  we observed similar results as in the experiment on the Rome graphs, that is, our unrestricted algorithm achieves best total resolution followed by the restricted variant and then the algorithm by Argyriou et al.; see Fig.~\ref{fig:northTotal-1}. For larger graphs, however, these three algorithms achieve similar results while still outperforming the algorithm by Huang et al. The results for the aspect ratio and number of crossings are similar to those of the crossing resolution experiment, with the exception of the fact that the algorithm of Huang et al. performs more stable with respect to the aspect ratio; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}. With respect to the number of iterations our two algorithms and the one by Argyriou et al. show similar behaviour needing more iterations than the algorithm by Huang et al.; see Fig.~\ref{fig:northTotal-4}. Observe that the number of iterations does not seem to correlate with the number of vertices.

@@

In the angular resolution experiments we again obtain a not-so-clear picture concerning the ranking of the algorithms, especially for higher number of vertices the ranking varies; see Fig.~\ref{fig:northAngular-1}. Only the algorithm by Huang et al. seems to be mostly at the last rank. Concerning the aspect ratio we see very good behaviour for our restricted variant and the algorithm by Huang et al. while the remaining two algorithms show large variance and much worse values; see Fig.~\ref{fig:northAngular-2}. For the number of crossings, we again observe that all algorithms achieve similar values, however, our unrestricted algorithm and the algorithm by Argyriou et al. achieve slightly higher values for larger graphs; see Fig.~\ref{fig:northAngular-3}. Finally, both our algorithms and the one by Huang et al. need a similar number of iterations for convergence which is lower than the one by Argyriou et al.; see Fig.~\ref{fig:northAngular-4}.

Summarizing this section, we conclude that compared to the Rome graphs the AT\&T graphs show a much higher variance regarding the various resolution measures. The crossing resolution results of our method still outperform the other approaches by far but especially for larger graphs as well as angular and total resolution the methods tend to perform similarly.

@@

We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the last year's contest winner. On a few graphs (namely, graphs 10 and 11), we even achieve better results. With a single exception (namely, graph 4), we easily outperformed our results from last year.
If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as last year's contest winner. We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parametrization will be needed to compute competitive solutions within the provided time.

>>>>

@@

\subfloat[\label{fig:crossingResolutionCrossingNumberGrid}{No.~of crossings vs no.~of vertices}]{
\centering
\scalebox{0.95}{
\centering
%======================================
% Crossing Resolution Crossing Number
%======================================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
xmin=10,
xmax=100,
ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
\addplot table [x=n, y=Crossing number rm1M, col sep=semicolon]{gridExperiment.csv};
\addplot table [x=n, y=Crossing number rm10k, col sep=semicolon] {gridExperiment.csv};
\addplot table [x=n, y=Crossing number rm1000, col sep=semicolon]{gridExperiment.csv};
\addplot table [x=n, y=Crossing number rm100, col sep=semicolon] {gridExperiment.csv};
\addplot table [x=n, y=Crossing number rm, col sep=semicolon] {gridExperiment.csv};
\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
\end{axis}
\end{tikzpicture}}}
\subfloat[\label{fig:crossingResolutionTimeGrid}{Iterations vs no.~of vertices}]{
\centering
\scalebox{0.95}{
\centering
%==================================
% Crossing Resolution Iterations
%==================================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
xmin=10,
xmax=100,
ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
ylabel={Iterations},
xlabel={Number of Vertices},ymajorgrids]
\addplot table [x=n, y=Iterations rm1M, col sep=semicolon]{gridExperiment.csv};
\addplot table [x=n, y=Iterations rm10k, col sep=semicolon] {gridExperiment.csv};
\addplot table [x=n, y=Iterations rm1000, col sep=semicolon]{gridExperiment.csv};
\addplot table [x=n, y=Iterations rm100, col sep=semicolon] {gridExperiment.csv};
\addplot table [x=n, y=Iterations rm, col sep=semicolon] {gridExperiment.csv};
\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
\end{axis}
\end{tikzpicture}}}
\caption{Our experimental results on the crossing resolution with different grid restrictions.
The double line corresponds to our unrestricted algorithm.}
\label{fig:experimentsGrid}
\end{figure}

In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates. In particular, we were interesting to see how the different quality measures that we evaluated in Section~\ref{sec:experiments} are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
%
\begin{inparaenum}[(i)]
\item $10^6 \times 10^6$
\item $10^4 \times 10^4$
\item $10^3 \times 10^3$, and
\item $10^2 \times 10^2$.
\end{inparaenum}
%
The test suite for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}. However, since our algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap. The layouts for each of the different sizes of the grid were computed with the variant of our algorithm that optimizes the crossing resolution; Fig.~\ref{fig:experimentsGrid} summarizes the results.

Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}. More precisely, a grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs. As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs. Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when grid size was set to $10^3 \times 10^3$.
%Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
%For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
For grid size $10^4 \times 10^4$, the drawings produced by our algorithm were on average by only $10^\circ$ worse than those produced by the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}), while the gap was closer for grid size $10^6 \times 10^6$.

The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}. The fact that aspect ratio of these drawings was worse can also explained of course by the fact that in most cases an improved drawing could not be reported.

As expected, the smaller the underlying grid is, the more crossings the computed drawings contain; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}. As a result, the unrestricted variant of our algorithm clearly outperforms all other ones. It is worth noting that the differences are clear between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$. Notably, there is only a slight improvement (in terms of the number of crossings) from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$. %This can partially be explained with the different choice of the initial drawing.
%
On the other hand, the number of iterations needed for convergence increases with the grid size; see Fig.~\ref{fig:crossingResolutionTimeGrid}, with the exception of the grid of size $10^2 \times 10^2$, which clearly indicates our previous observation that for the vast majority of the graphs an improved drawing could not be reported.

In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid, as long as the grid is not too small. However, the computation of a grid drawing takes longer, which is of course expected.
%than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order
Finally, note that the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution) but also with respect to the number of iterations needed to converge.
%To improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution. Further experiments in this direction are needed.

@@

In this section, we report the results of our expertimental evaluation (on the crossing, total and angular resolutions) for the non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}). Note that we did not impose any  grid constraints our algorithms. The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}. In general we observed that the variance of results is much larger than in the experiments on the Rome graphs. This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of the graphs in this second test set varies vastly between different graph sizes.

@@

For the crossing resolution we observed that both variants of our algorithm again outperformed the two force-directed algorithms; see Fig.~\ref{fig:northCrossing-1}. Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution, as the curves are nearly parallel. By all these results, we can classify the graphs into ``hard'' or ``easy'' when maximizing their crossing resolution. In particular, graphs with 50 to 70 vertices appear to be harder to improve than graphs with 70 to 80 vertices. Regarding the aspect ratio of the produced drawings, we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force-directed algorithms appears to be unstable resulting in a large variance. Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio, which is much lower than the one of our unrestricted algorithm for larger graphs. All four algorithms behave nearly the same in terms of the number of crossings; see Fig.~\ref{fig:northCrossing-3}. In terms of the number of iterations, we observe that somewhat surprisingly the algorithm of Argyriou et al. converges in the least amount of iterations, while the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.

\begin{figure}[tb]
\centering
\subfloat[\label{fig:northTotal-1}{Total resolution vs no.~of vertices}]{
\centering
\scalebox{.95}{
\centering
%======================
% Total Resolution
%======================
\begin{tikzpicture}
\begin{axis}[
axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
width=0.475\textwidth,
cycle list name=mylist,
mark repeat={5},xlabel={Number of Vertices},ylabel={Total Resolution},
xtick={10,20,30,40,50,60,70,80,90,100},
xmin=10,
xmax=100,
ytick={10,20,30,40,50,60,70,80,90},
tick pos=left,ymajorgrids]
\addplot table [x=avg, y=tr rm, col sep=semicolon]{totalResolutionNorth.csv};
\addplot table [x=avg, y=tr rm-1, col sep=semicolon] {totalResolutionNorth.csv};
\addplot table [x=avg, y=tr fa, col sep=semicolon]{totalResolutionNorth.csv};
\addplot table [x=avg, y=tr tr, col sep=semicolon] {totalResolutionNorth.csv};
\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
\end{axis}
\end{tikzpicture}}}
\subfloat[\label{fig:northTotal-2}{Aspect ratio vs no.~of vertices}]{
\centering
\scalebox{.95}{
\centering
%===================================
% Total Resolution Aspect Ratio
%===================================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
xlabel={Number of Vertices},
xmin=10,
xmax=100,
ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
\addplot table [x=avg, y=ar rm, col sep=semicolon]{totalResolutionNorth.csv};
\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {totalResolutionNorth.csv};
\addplot table [x=avg, y=ar fa, col sep=semicolon]{totalResolutionNorth.csv};
\addplot table [x=avg, y=ar tr, col sep=semicolon] {totalResolutionNorth.csv};
\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
\end{axis}
\end{tikzpicture}}}

@@

In the total resolution experiment, we observed similar results as in the experiment on the Rome graphs for small graphs, that is, our unrestricted algorithm outperforms the other three ones, while the restricted variant of our algorithm yields drawings and then the algorithm by Argyriou et al.; see Fig.~\ref{fig:northTotal-1}. For larger graphs, however, these three algorithms achieve similar results while still outperforming the algorithm by Huang et al. The results for the aspect ratio and number of crossings are similar to those of the crossing resolution experiment, with the exception of the fact that the algorithm of Huang et al. performs more stable with respect to the aspect ratio; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}. With respect to the number of iterations our two algorithms and the one by Argyriou et al. show similar behaviour needing more iterations than the algorithm by Huang et al.; see Fig.~\ref{fig:northTotal-4}. Observe that the number of iterations does not seem to correlate with the number of vertices.

@@

In the angular resolution experiment, we again obtain a not-so-clear picture concerning the ranking of the algorithms, especially for higher number of vertices the ranking varies; see Fig.~\ref{fig:northAngular-1}. Only the algorithm by Huang et al. seems to be mostly at the last rank. Concerning the aspect ratio we see very good behaviour for our restricted variant and the algorithm by Huang et al. while the remaining two algorithms show large variance and much worse values; see Fig.~\ref{fig:northAngular-2}. For the number of crossings, we again observe that all algorithms achieve similar values, however, our unrestricted algorithm and the algorithm by Argyriou et al. achieve slightly higher values for larger graphs; see Fig.~\ref{fig:northAngular-3}. Finally, both our algorithms and the one by Huang et al. need a similar number of iterations for convergence which is lower than the one by Argyriou et al.; see Fig.~\ref{fig:northAngular-4}.

Summarizing, we conclude that compared to the Rome graphs, the AT\&T graphs show a much higher variance regarding the various resolution measures. %The crossing resolution results of our method still outperform the other approaches by far but especially for larger graphs as well as angular and total resolution the methods tend to perform similarly.

@@

We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the last year's contest winner. On a few graphs (namely, graphs 10 and 11), we even achieve better results. With a single exception (namely, graph 4), we easily outperformed our results from last year.
If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as last year's contest winner. We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parameterization will be needed to compute competitive solutions within the provided time.
